Manuscript

### 2025-03-22 Manuscript_MD_ScoreReLiability.qmd
added bullet point rant

>>>>

-   examing heterogeneity in standardized mean differences before and
    after correction for differences in score reliability using an
    attenuation correction.

-   Find that a) in the majority of cases, the change in ES
    heterogeneity is small to non-existent, implying that differences in
    score reliability do not serve an appropriate explanation of ES
    heterogeneity.

-   Find that b) in the majority of cases, the direction of change in ES
    heterogeneity does not align with assumptions formulated in W&D and
    H&S. In close to all cases (all cases if we only look at
    statistically significant differences in ES), the attenuation
    correction leads to increases estimates of ES heterogeneity. This
    implies that W&D and H&S generally misunderstand how score
    reliability heterogeneity affects ESs.

-   Using analytical arguments, we derived a general rule on under which
    conditions we can expect an attenuation correction to reduce ES
    heterogeneity. These conditions depend on how both the mean value
    and the heterogeneity of score variance is reduced by correction for
    sore reliability. At the very least, $R_1^3 < R_2$. This means that
    as long as the reduction in score variance heterogeneity is larger
    than the reduction in mean score variance, ES heterogeneity will
    grow as we follow the attenuation correction procedure.

-   We know what an acceptable average score reliability looks like in
    the area of Psychology. Researchers largely target a score
    reliability of at least .7, preferably .8. An average score
    reliability of .7 implies that 70% of the variance observed in
    scores can actually be attributed to true underlying differences.
    However, from our derived rule, we know that at an average score
    reliability of .7, the heterogeneity in score variances across
    administration sites that can be attributed to differences in the
    true score variance $\sigma^2_T$ may be at most .343 or 34.3%.
    Accordingly, at least 65.7%, almost two thirds of the score variance
    heterogeneity, needs to be due to differences in error score
    variance, differences in the measurement precision.

-   However, in the examples we re-analysed throughout this text, as
    well as the majority of research studies done in the behavioural
    sciences, no proper sampling techniques are employed. Most samples
    where these instruments are applied come to be through convenience
    sampling, where most often students self-select into studies.
    Without a valid sampling scheme, we have essentially have no basis
    to assume anything about the true score variance. Most importantly,
    however, we have no reason to assume that the true score variance
    could potentially be homogeneous or stable across administration
    sites. Instead, it is very likely that the convenience sampling
    procedures produce samples that vary extremely strongly in how they
    are put together, leading to large degrees of true score variance
    heterogeneity. At the same time, most researchers are aware and
    trained in assessing score reliability.

-   Taken together, these facets lead to what we observed in the
    phenomena examined in this text: Score reliability in general is
    acceptable, leading to a small, but constant reduction in mean score
    variance if we apply an attenuation correction. At the same time,
    there are substantial differences in true score variance, implying
    that most of the score variance heterogeneity can not be attributed
    to differences in error score variance. Since only differences in
    error score variance can be corrected for, the attenuation
    correction has a very small effect on heterogeneity estimates in
    score variance. Therefore, as indicated by our rule, an attenuation
    correction in these setting can not deliver a reduction in ES
    heterogeneity. This also means that, in such settings, score
    reliability does not explain any ES heterogeneity, if anything, it
    masks it to a small degree.

-   If we, in the space of behavioural sciences, keep making use of
    convenience samples simply because they are so convenient, we can
    not expect anything to change. Most likely, ES heterogeneity will
    remain a substantial factor hampering the interpretability of
    original and replication studies. We could learn more about
    standardized effect sizes and why our phenomena vary across
    settings, populations or continents by not following the current
    status quo of convenience samples. Until this status quo changes,
    however, we will keep wasting valuable resources by attempts of
    discussing ES heterogeneity or assessing uninformative differences
    in score reliability.
